---
title: "Basic R training"
author: "Anil Adhikari"
date: "3/13/2022"
output:
  html_document: default
  pdf_document: default
---
# Data input in R
There are three different ways of data input into R.

- Use RStudio features - Method 1.  
File > Open file > [select the file you want to input]

- Use RStudio features - Method 2.  
File > Import Dataset > From CSV / From Excel > specify input features

- Using R commands.  
Naviagte to the directory with data files and use appropriate commands.  


```{r}
# check your working directory
getwd()

# set your working directory to the folder with data files
setwd("/Users/aniladhikari/Desktop/RLab")

# View the file names
list.files()

# Use correct file name to run input commands
dat1 <- read.csv("Wheatc.csv")

# For text files
dat2 <- read.table("Wheat_91.txt", sep = "\t", header = TRUE)
```

# Data exploratory analysis 
1. Check if the data is in correct format. For example, check for typos, alphabets in place of numbers, missing row column names, missing data etc.

2. Check the distribution of data for outliers. There can be two types of outliers, true outliers, outliers due to inaccuracy in data collection, input, unit differences etc. We should try to see if there are any outliers of the second kind and rectify them.

```{r}
# checking data for correct data type and format
head(dat1)

# check data types in each columns
str(dat1)

# check for missing data
table(is.na(dat1))
```

```{r}
# checking for data accuracy

# basic histogram for distribution
hist(dat1$yield, main = "Yield ~ variety", xlab = "yield")
```
```{r}
# create a boxplot
library(ggplot2)
ggplot(dat1, aes(x=variety, y=yield, fill = variety)) +
  geom_boxplot()
```
# Analysis of variance  
```{r}
model1 <- lm(yield ~ variety, data = dat1)
anova(model1)
```
# check model fit
```{r}
plot(model1)
```
# mean separation
```{r}
install.packages("agricolae")
library("agricolae")

# Tukey's HSD
HSD.test(model1,"variety", alpha = 0.05, group = TRUE)

```

